One-armed spirals in locally isothermal, radially structured self-gravitating discs

### 2015-01-13 paper/paper.tex
almost done with first revision

% planet is popular answer for spiral structure
The attractive explanation for asymmetries in circumstellar discs is
disc-planet interaction. In particular, spiral structures  
naturally arise from the gravitational interaction between a
planet and the gaseous protoplanetary disc it is embedded in
\citep[see, e.g.][for a recent review]{baruteau13b}. Thus, the
presence of spiral arms in circumstellar discs could be signposts of 
planet formation.

@@

{\bf
  In this work we identify a new mechanism that leads to the growth of
  one-armed spirals in astrophysical discs. We show
  that when the disc temperature is prescribed (called locally isothermal
  discs), the usual statement of the conservation of angular momentum
  for linear perturbations acquires a source term proportional to the
  temperature gradient. This permits angular momenum exchange between
  linear perturbations and the background disc. This `background
  torque' may act to destabilise low-frequency non-axisymmetric
  trailing waves when the disc temperature decreases outwards.       
}

@@

{\bf
\section*{Acknowledgments}
I thank K. Kratter and A. Youdin for valuable discussions, and the
anonymous referee for comments that significantly improved this paper. All
computations were performed on the El Gato cluster at the University
of Arizona. This material is based upon work supported by the National
Science Foundation under Grant No. 1228509. 
}
\bibliographystyle{mn2e}
\bibliography{ref}

>>>>

% planet is popular answer for spiral structure
The attractive explanation for asymmetries in circumstellar discs is
disc-planet interaction. In particular, spiral structures  
naturally arise from the gravitational interaction between a
planet and the gaseous protoplanetary disc it is embedded in
\citep[see, e.g.][for a recent review]{baruteau13b}. Thus, the
presence of spiral arms in circumstellar discs could be signposts of 
planet formation \citep{juhasz14}.

@@

{\bf
  In this work we identify a new mechanism that leads to the growth of
  one-armed spirals in astrophysical discs. We show
  that when the disc temperature is prescribed (called locally isothermal
  discs), the usual statement of the conservation of angular momentum
  for linear perturbations acquires a source term proportional to the
  temperature gradient. This permits angular momenum exchange between
  linear perturbations and the background disc. This `background
  torque' can destabilise low-frequency non-axisymmetric
  trailing waves when the disc temperature decreases outwards.       
}

@@

{\bf
\section*{Acknowledgments}
I thank K. Kratter, Y. Wu and A. Youdin for valuable discussions, and the
anonymous referee for comments that significantly improved this paper. All
computations were performed on the El Gato cluster at the University
of Arizona. This material is based upon work supported by the National
Science Foundation under Grant No. 1228509. 
}
\bibliographystyle{mn2e}
\bibliography{ref}
